---
title: "01.DownloadData"
author: "Sara Contu"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true 
    number_sections: true
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=FALSE, message=FALSE}
  # install and load packages

  library("dplyr")

 
```

 Load the datasets

1. Download and load PREDICTS 2016 release data (olddb hereafter). The data has been downloaded from the [NHM data portal 2016 release](https://data.nhm.ac.uk/dataset/the-2016-release-of-the-predicts-database).


```{r }

  olddb <- readRDS("Data/database1strelease.rds")


```
2. Download and load PREDICTS November 2022 release data (novdb hereafter). The data can be downloaded from the [NHM data portal](https://data.nhm.ac.uk/dataset/release-of-data-added-to-the-predicts-database-november-2022)

```{r }

  novdb <- readRDS("Data/database2ndrelease.rds")


```


## Explore the data 

 - source ID
 - number of sites (SSBS)
 - number of species (this is how species were counted for the first paper: How did you count 7,136 species? I make it 10,635 species. I used the `CountSpecies` function in yarg, which is what we used to count the number of species for the 2017 paper. It uses the following
  Taxa resolved to species or infraspecies count once per taxon name
  Higher ranks count once per name per study)
 - number and identities of countries (map?)

```{r, echo=FALSE}

  sourceID_olddb <- as.data.frame(unique(olddb$Source_ID)) # identities and length of source IDs 
  
  SSBS_olddb <- as.data.frame (unique(olddb$SSBS)) # this is wrong - need to fix it
  
  
 # species_olddb <-
  # I will need to use the function in yarg
  
  countries_olddb <- as.data.frame(unique(olddb$Country)) # identities and lenght of the countries included in the olddb
  

```

 
```{r , echo=FALSE}

  novdb <- readRDS("Data/database2ndrelease.rds")

  sourceID_novdb <- as.data.frame(unique(novdb$Source_ID)) # identities and length of source IDs 
  
  SSBS_novdb <- as.data.frame (unique(novdb$SSBS)) # this is wrong - need to fix it
  
 # species_olddb <-
  # I will need to use the function in yarg
  
  countries_novdb <- as.data.frame(unique(novdb$Country)) # identities and lenght of the countries included in the olddb
  

```
