---
title: "01.DownloadData"
author: "Sara Contu"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true 
    number_sections: true
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=FALSE, message=FALSE}
  # install and load packages

  library("dplyr")

 
```




# Source relevant data

This work is aimed at addressing:

"How much difference is made in model fit and projected BII by new data, LUH2 recuration and taxon-level random effect?"

Old data vs old + new data LUH1 (mod1 ~ mod2) {2016 MODELS}

Old data vs old + new data LUH2 (mod3 ~ mod4) {2016 MODELS}

Old data vs old + new data LUH1 (mod5 ~ mod6) {2016 MODELS + Taxon-fold}

Old data vs old + new data LUH2 (mod7 ~ mod8) {2016 MODELS + Taxon-fold}

Below I list the data and sources of data necessary for this analysis. 
N.B. This list will grow and change as i get more in the analysis. please feel free to add to the list.


## Database release 2016 with old LUH1
  -  Download and load PREDICTS 2016 release data (olddb hereafter) from the [NHM data portal 2016 release](https://data.nhm.ac.uk/dataset/the-2016-release-of-the-predicts-database).
  
  
```{r }

  olddb <- readRDS("Data/database1strelease.rds")


```
  

## Database release 2016 LUH2
  - need to ask Andy/Adriana 
  
## Database release 2022 LUH1 
  -Download and load PREDICTS November 2022 release data (novdb hereafter) from the [NHM data portal](https://data.nhm.ac.uk/dataset/release-of-data-added-to-the-predicts-database-november-2022)
```{r }

  novdb <- readRDS("Data/database2ndrelease.rds")


```
  
## Database release 2022 LUH2
   - need to ask Andy/Adriana 

## (Database release 2016 with new LUH1 (some studies may have been recurated - so check for any diff))
  - download the database extract from psql-predicts and filter for 2016 database release sources to compare difference in LUH1 with the old version)
  

# Explore the data - preparation for modelling

 - source ID
 - number of sites (SSBS)
 - number of species (N.B. the number of species for the first release were calculated using the `CountSpecies` function in yarg. It uses the following
  Taxa resolved to species or infraspecies count once per taxon name
  Higher ranks count once per name per study)
 - number and identities of countries (map?)

```{r, echo=FALSE}

  sourceID_olddb <- as.data.frame(unique(olddb$Source_ID)) # identities and length of source IDs 
  
  SSBS_olddb <- as.data.frame (unique(olddb$SSBS)) # this is wrong - need to fix it
  
  
 # species_olddb <-
  # I will need to use the function in yarg
  
  countries_olddb <- as.data.frame(unique(olddb$Country)) # identities and lenght of the countries included in the olddb
  

```

 
```{r , echo=FALSE}

  novdb <- readRDS("Data/database2ndrelease.rds")

  sourceID_novdb <- as.data.frame(unique(novdb$Source_ID)) # identities and length of source IDs 
  
  SSBS_novdb <- as.data.frame (unique(novdb$SSBS)) # this is wrong - need to fix it
  
 # species_olddb <-
  # I will need to use the function in yarg
  
  countries_novdb <- as.data.frame(unique(novdb$Country)) # identities and lenght of the countries included in the olddb
  

```


